\begin{appendices}

\addtocontents{toc}{\protect\renewcommand{\protect\cftchappresnum}{\appendixname\space}}
\addtocontents{toc}{\protect\renewcommand{\protect\cftchapnumwidth}{7em}}

\chapter{Artificial Intelligence Assistance Disclosure}

In accordance with academic integrity standards and emerging best practices for transparent research documentation, this appendix discloses the use of artificial intelligence tools in the preparation of this dissertation proposal. Such disclosure reflects the author's commitment to scholarly transparency and recognition of evolving norms within the academic community.

\section{AI Tools Utilized}

This dissertation proposal utilized Claude Opus 4.5 \& Sonnet 4.5, developed by Anthropic, to assist with specific aspects of document preparation. The AI tool was employed under the direct supervision of the author, with all outputs subjected to critical review, verification, and revision as necessary to ensure accuracy, appropriateness, and alignment with the research objectives. The author exercised editorial judgment over all AI-assisted contributions.

\section{Scope of AI Assistance}

AI assistance was limited to the following editorial and mechanical activities:

\begin{itemize}
    \item Grammar and spelling review to identify and correct mechanical errors
    \item Sentence structure refinement to improve clarity and readability
    \item Organizational suggestions to enhance document flow and logical progression
    \item Formatting consistency verification across sections and chapters
    \item Citation format verification for compliance with IEEE standards
\end{itemize}

\section{Scope Limitations}

AI tools were explicitly excluded from the following substantive activities:

\begin{itemize}
    \item Generation of original research ideas, hypotheses, or theoretical frameworks
    \item Literature search, source identification, or citation selection
    \item Data analysis, statistical calculations, or interpretation of results
    \item Writing of substantive original content without subsequent author revision
    \item Determination of research methodology or survey instrument design
\end{itemize}

The intellectual contributions of this research---the identification of research gaps, the development of the theoretical framework, the design of the survey instrument, and the interpretation of findings---remain solely the work of the author.

\section{Author Responsibility}

The author maintains full responsibility for all content, arguments, analyses, and conclusions presented in this dissertation. All AI-assisted text was critically reviewed and revised by the author to ensure it accurately represents the author's original thinking and research contributions. The employment of AI tools for editorial assistance does not diminish the author's intellectual ownership of this work, nor does it transfer scholarly credit for the ideas and arguments herein.

\section{Rationale for Disclosure}

This disclosure is provided in the spirit of transparency and in recognition of evolving academic standards regarding AI tool usage. As AI capabilities continue to advance, clear documentation of how these tools are employed in academic work supports research integrity and enables appropriate evaluation of scholarly contributions. The author believes that transparent disclosure serves the academic community better than silent utilization, and that establishing norms for responsible AI assistance in scholarly work represents an important contribution to academic discourse.

\end{appendices}