\chapter*{\Large \uppercase{Acknowledgments}}
\chaptermark{Acknowledgments}\label{chap:Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

I extend my profound gratitude to my dissertation committee for their guidance, patience, and insight throughout this journey. Their expertise and steadfast encouragement proved instrumental in bringing this work to fruition.

To my grandparents---Milt and Norma Hoag, Jim and Bertha Bonar---whose legacy of perseverance and principled living continues to inspire me. The values you instilled endure.

To my wife, Sarah, who has been my harbor and my calm through the many years of balancing academic pursuits with professional responsibilities. Your unwavering support made this achievement possible. No words suffice.

To my father, Randy, my mother, Rita, and my brother, Joe. Though I was not always the easiest child to raise, you never wavered in guiding me toward the right path. Your belief in me has been a constant source of strength, and I carry it with me still.

To my cousin, Doctor Kathryn Fishman-Weaver, whose academic achievements set a standard of excellence and inspired the rest of our family to reach higher. Your accomplishments demonstrated what dedication and intellect can achieve.

Finally, I wish to acknowledge the mentors whose wisdom has shaped my professional journey: Keith Summerson, Dale Kurth, Micah Mogle, Kavi Parupally, Doctor John Hastings, Doctor Kyle Cronin, Doctor Casey Mayfield, Kelly Ortberg, Greg Kouski, and Charles Espy Jr. Though years may have passed since our paths crossed, your words of encouragement and guidance continue to resonate.
